
---
title: "Simple document"
author: "Catherine Mauro"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

 
# Problem 1

First, we should load the **tidyverse** into our code.

```{r}

library(tidyverse)

```

Next, we can upload our **Instacart** data using the *p8105.datasets* code provided on the website.

```{r}

library(p8105.datasets)
data("instacart")

```


After loading the set, we can observe some details regarding the data:

There are `r nrow(instacart)` rows and `r ncol(instacart)` columns in **instacart**.

The variables include `r colnames(instacart)`. Columns like order, product, and user id record unique numeric identifiers, while product name and department record descriptive character observations from the
data.


Nest, we can write some code to find more about our **instacart** data.

```{r}

instacart %>%
  count(aisle) %>%
  arrange(desc(n))

```


There are `r length(unique(pull(instacart, aisle)))` aisles. The aisles with the 
most items include _fresh vegetables_, _fresh fruits_, _packaged vegetables fruits_, 
_yoghurt_, and _packaged cheese_, all which have more than **41,000** observations.


Next, we can make a plot to show the number of items ordered in each aisle for
those aisles with more than **10,000** observations.


```{r}

instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
    mutate(
      is.factor(aisle),
      fct_reorder(aisle, n, desc)
    ) %>%
  ggplot(aes(x = aisle, y = n)) + geom_point() + theme(axis.text.x = element_text
  (angle = 90, hjust = 1, vjust = 3)) +
  labs(
    title = "Aisle Plot",
    x = "Aisle Name",
    y = "Number of Items Ordered",
    caption = "data from p8105.datasets"
  )
```


After arranging a graph for all of the aisles with over 10000 items ordered,
we can make a table further exploring popular items for _baking ingredients_,
 _dog food care_, _packaged vegetables fruits_.

```{r}

instacart %>%
  group_by(aisle) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  count(product_name) %>%
  mutate(product_rank_in_aisle = min_rank(desc(n))) %>%
  filter(product_rank_in_aisle <= 3) %>%
  arrange(aisle, product_rank_in_aisle) %>%
  knitr::kable()
```

Finally for this data set, we can create a 2x7 table tracking the mean hour of day at which certain apples
and ice creams are ordered each day of the week.


```{r}

instacart %>%
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  summarize(mean_hod = round(mean(order_hour_of_day), 2)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hod
  )



```










